Computing in the Field of Rationals

### 2012-06-19 rational_ext.tex
_

\authorinfo{Roshan P. James}
           {Indiana University}
           {rpjames@indiana.edu}
\authorinfo{Amr Sabry}
           {Indiana University}
           {sabry@indiana.edu}

@@

  Andrzej Filinski's masters thesis
  \cite{Filinski:1989:DCI:648332.755574} suggested a remarkable
  symmetry underlying computation in the \lcal. Filinski showed that
  values are dual to continuations and that functions are dual to
  delimited continuations, thereby suggesting that
  values-and-functions and continuations-and-delimited-continuations
  are mirror images of the same phenomena.

  Previous work on \emph{information effects} established the notion
  of information preserving computation
  \cite{infeffects,rc2011,rc2012} and the core calculus for computing
  with the isomorphisms of finite types, {{Pi}}. Given the
  first-order nature of {{Pi}} however it was unclear how one may
  express first-class functions and duality in the sense of Filinski.

  We present a computational model whose types are \emph{the field of
    rational numbers}. This computational model is derived
  systemmatically from the {{Pi}} by considering a symmetric
  notion of duality. Unlike the \lcal which shows a classical demorgan
  duality, here we have two axis of dualization and hence two
  dualities -- an additive duality, namely negative types and a
  multiplicative duality, namely fractional types.

>>>>

\authorinfo{Nicolas Bourbaki}

%% \authorinfo{Roshan P. James}
%%            {Indiana University}
%%            {rpjames@indiana.edu}
%% \authorinfo{Amr Sabry}
%%            {Indiana University}
%%            {sabry@indiana.edu}

@@

  This paper connects the line of work starting with
  Filinski\cite{Filinski:1989:DCI:648332.755574} on the duality of
  computation \cite{Curien:2000,DBLP:conf/rta/Wadler05} with work on
  information preserving computation \cite{infeffects,rc2011}.
  Andrzej Filinski's masters thesis
  \cite{Filinski:1989:DCI:648332.755574} suggested a remarkable
  symmetry underlying computation in the \lcal. Filinski showed that
  values are dual to continuations and that functions are dual to
  delimited continuations, thereby suggesting that
  values-and-functions and continuations-and-delimited-continuations
  are mirror images of the same phenomena.

  Previous work on \emph{information effects} established the notion
  of information preserving computation \cite{infeffects} and the core
  calculus for computing with the isomorphisms of finite types,
  {{Pi}}. Given the first-order nature of {{Pi}} however it was
  unclear how one may express first-class functions and duality in the
  sense of Filinski.

  We present a computational model whose types are \emph{the field of
    rational numbers}. This computational model is derived
  systemmatically from the {{Pi}} by considering a symmetric notion of
  duality. Unlike the \lcal which shows a classical De Morgan duality,
  here we have two axis of dualization and hence two dualities -- an
  additive duality, namely negative types and a multiplicative
  duality, namely fractional types.
